---
title: "Used Cars Dataset (Kaggle)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("cars.R")
```

## Cars in Georgia:

### Scatter Plot of Cars Sold vs Average Price:

```{r}
df <- tibble(make_count_GA)
ggplot(df, aes(`Number Sold`, `Average Price`)) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_jitter() +
  xlab("Number Sold") +
  ylab("Average Price") +
  ggtitle("Number Sold vs Average Price") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth() #using method 'loess'
```

### Box Plot

#### Check Outliers

```{r}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(make_count_GA$`Number Sold`, main="Number Sold") 
boxplot(make_count_GA$`Average Price`, main="Average Price")
```

### Density Plot of Average Car Price:

#### Checking Normality

```{r}
plot(density(make_count_GA$`Average Price`), main="Density Plot: Average Price", ylab="Frequency", sub=paste("Skewness:", round(e1071::skewness(make_count_GA$`Average Price`), 2))) 
polygon(density(make_count_GA$`Average Price`), col="cornflowerblue")
```

### Build Linear Model (x: Number Sold, y: Average Price)

```{r}
linearMod <- lm(make_count_GA$`Average Price` ~ make_count_GA$`Number Sold`, data = cars_in_GA)
summary(linearMod)
```


## Comparing Transmission Types and Prices:

### Manual vs Auto:

#### Density Plot

```{r}
options(scipen=5)
df_m <- tibble(manual_trans)
df_a <- tibble(auto_trans)

plot_m <- ggplot(df_m, aes(Price), col="cornflowerblue", fill = "cornflowerblue") +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_density(kernel = "gaussian") + xlab("Price") +
  ggtitle("Density Plot of \n Manual Cars' Price") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_density(color="darkblue", fill="lightblue") +
  geom_vline(aes(xintercept=mean(Price)),
            color="blue", linetype="dashed", size=1)

plot_a <- ggplot(df_a, aes(Price)) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_density(kernel = "gaussian") +
  xlab("Price") +
  ggtitle("Density Plot of \n Automatic Cars' Price") +
  theme(plot.title = element_text(hjust = 0.5))  +
  geom_density(color="darkblue", fill="lightblue") +
  geom_vline(aes(xintercept = mean(Price)),
            color="blue", linetype="dashed", size=1)

summary(auto_trans$Price)
summary(manual_trans$Price)
plot_grid(plot_m, plot_a)
```

#### Box Plot

```{r}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(manual_trans$Price, main="Manual Transmission Car Prices") 
boxplot(auto_trans$Price, main="Automatic Transmission Car Prices")
```

### Takeaway


There are way more automatic transmission cars than manual transmission. The prices of manual transmission cars are approximately normally distributed, with most cars being in the $12,000 to $20,000 price range. On the other hand, automatic cars have a very positively skewed distribution. Most cars are within $80,000, but there are many outliers -- cars priced up to almost $500,000. Despite this, I believe we **cannot** exclude these outliers as the high amount of expensive vehicles could be related to state mean income. Further exploration on this topic is below.



## Exploring Relationship with State Mean Income and Mean Price of Cars Sold in Each State:

### Mean Income by State

#### Scatterplot
```{r}
df <- tibble(state_mean_income)
ggplot(df, aes(States, `Mean Income`)) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_jitter() +
  xlab("State Name") +
  ylab("Mean Income") +
  ggtitle("States and Their Mean Income") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Density Plot

```{r}
ggplot(df, aes(`Mean Income`)) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_density(kernel = "gaussian") +
  xlab("Mean Income") +
  ggtitle("Density Plot of \n State Mean Income") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_density(color="darkblue", fill="lightblue") +
  geom_vline(aes(xintercept=mean(`Mean Income`)), color="blue", linetype="dashed", size=1)
```

### Mean Price of Cars Sold in Each State

#### Scatterplot

```{r}
#selecting and cleaning data
price_and_state <- test %>%
  select(Price, State) %>%
  mutate(State = toupper(State))%>%
  group_by(State) %>%
  mutate(Price = mean(Price)) %>%
  unique() 

#sort by state name
price_and_state <- price_and_state[order(price_and_state$State), ]

#plot
df_ps <- tibble(price_and_state)
ggplot(df_ps, aes(State, Price)) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_jitter() +
  xlab("State Name") +
  ylab("Average Car Price") +
  ggtitle("States and Average Car Price") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Density Plot

```{r}
ggplot(df_ps, aes(Price)) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_density(kernel = "gaussian") +
  xlab("Price") +
  ggtitle("Density Plot of \n Average Car Price") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_density(color="darkblue", fill="lightblue") +
  geom_vline(aes(xintercept=mean(Price)), color="blue", linetype="dashed", size=1)
```


### Combined Data (State, Mean Income, Mean Price of Cars)

```{r}
state_mean_income <- state_mean_income[state_mean_income$States != "Puerto Rico", ]
state_price_income <- tibble(price_and_state, state_mean_income) 
state_price_income <- state_price_income %>%
  select(State, Price, `Mean Income`)
summary(state_price_income)
```

#### 2-Variable Scatter Plot

```{r}
df_ps <- tibble(state_price_income)
ggplot(df_ps, aes(`Mean Income`, Price)) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_jitter() +
  geom_smooth() +
  xlab("Mean Income") +
  ylab("Average Car Price") +
  ggtitle("States' Mean Income vs Average Car Price") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Construct Regression Model

##### Linear

```{r}
model_l <- lm(Price ~ `Mean Income`, data = state_price_income)
summary(model_l)
```


##### Polynomial

```{r}
model_p5 <- lm(Price ~ poly(`Mean Income`, 5), data = state_price_income)
model_p10 <- lm(Price ~ poly(`Mean Income`, 10), data = state_price_income)
model_p15 <- lm(Price ~ poly(`Mean Income`, 15), data = state_price_income)
model_p20 <- lm(Price ~ poly(`Mean Income`, 20), data = state_price_income)


summary(model_p5)
summary(model_p10)
summary(model_p15)
summary(model_p20)


```

#### Conclusion

There seems to be no significant relationship between the average income and average price of cars sold.



